### 2019-10-31 Manuscript/manuscript.tex
add note about polarity chosen in Superior-Slave figure
add details and Nimmo2000a citation regarding a dynamo implying plate tectonics

%The poles on this figure should be labeled
\begin{figure}
\centering
\includegraphics[width=\textwidth]{../Figures/Superior_Slave_reconstructions.pdf}
\caption{\small{\textbf{Paleogeographic reconstructions developed using poles from the Superior, Slave and Rae provinces. Paleomagnetic poles are shown colored to match their respective province with these provinces shown in present-day coordinates and labeled in the 0 Ma panel. Poles with ages that are within 25 million years of the given time slice are shown. The relatively well-resolved pole paths from the Superior and Slave provinces (Fig. \ref{fig:Laurentia_poles}) that are utilized for these reconstructions provide strong support for differential plate tectonic motion between 2220 and 1850 Ma.}}}
\label{fig:Superior_Slave_recons}
\end{figure}

@@

\subsection{Conclusion}

@@

Another perspective on Proterozoic tectonics, is that the record is one of intermittent subduction \citep{Silver2008a, ONeill2013a}. In such a model, there are extended intervals with a stagnant lid alternating with intervals of differential plate motion. In particular, it was been argued that the Mesoproterozoic Era (1.6 to 1.0 Ga) is an interval when Earth was in a stagnant regime \citep{Silver2008a, ONeill2013a}. The accretionary history of Laurentia following the amalgamation of the Archean provinces is difficult to reconcile with such an interpretation (Figs. \ref{fig:Laurentia_map} and \ref{fig:tectonic_history}). An additional constraint comes from the presence of reliable paleomagnetic poles themselves that are complete with baked contact tests. In a stagnant lid regime, there would not be sufficient heat flow to sustain a geodynamo and Earth should not have had a geomagnetic field that would have enable the preservation of paleomagnetic poles with positive field tests. The record of these poles also show that there was motion of Laurentia (Figs. \ref{fig:Laurentia_paleolatitude} and \ref{fig:Laurentia_reconstructions}). Using data from Laurentia alone, however, it is difficult to ascertain whether this motion is due to plate tectonic motion or rotation of the entire solid Earth through true polar wander. True polar wander can lead to changing position relative to the spin axis even with a stagnant lid. One interval when the Laurentian paleomagnetic record demands that some of the motion is through differential plate tectonics is in the latest Mesoproterozoic. At that time, the pole path is very well-resolved with many high-quality paleomagnetic poles between 1110 and 1070 Ma (Table 2; \ref{fig:Laurentia_poles}). The progression of the poles requires rotation about an Euler pole that is distinct from a great circle path which would result if the motion were solely due to true polar wander \citep{Swanson-Hysell2019a}. These poles constrain rapid motion of Laurentia leading up to collisional orogenesis associated with the Grenvillian orogeny as illustrated in Figure \ref{fig:Grenville_reconstructions}. These data provide strong evidence for differential plate motion at the time and are inconsistent with a stagnant lid. Rather the orogenic cycle of the Mesoproterozoic bears similarity with that of the Paleozoic and reveals Laurentia to have been a central player in the building of amalgamated continents associated with Rodinia and Pangea.

>>>>

%The poles on this figure should be labeled
\begin{figure}
\centering
\includegraphics[width=\textwidth]{../Figures/Superior_Slave_reconstructions.pdf}
\caption{\small{\textbf{Paleogeographic reconstructions developed using poles from the Superior, Slave and Rae provinces. The polarity options that are chosen for the provinces are those that minimize total apparent polar wander path length. Paleomagnetic poles are shown colored to match their respective province with these provinces shown in present-day coordinates and labeled in the 0 Ma panel. Poles with ages that are within 25 million years of the given time slice are shown. The relatively well-resolved pole paths from the Superior and Slave provinces (Fig. \ref{fig:Laurentia_poles}) that are utilized for these reconstructions provide strong support for differential plate tectonic motion between 2220 and 1850 Ma.}}}
\label{fig:Superior_Slave_recons}
\end{figure}

@@

\subsection{The record implies plate tectonics throughout the Proterozoic}

@@

Another perspective on Proterozoic tectonics, is that the record is one of intermittent subduction \citep{Silver2008a, ONeill2013a}. In such a model, there are extended intervals with a stagnant lid alternating with intervals of differential plate motion. In particular, it was been argued that the Mesoproterozoic Era (1.6 to 1.0 Ga) is an interval when Earth was in a stagnant regime \citep{Silver2008a, ONeill2013a}. The accretionary history of Laurentia following the amalgamation of the Archean provinces is difficult to reconcile with such an interpretation (Figs. \ref{fig:Laurentia_map} and \ref{fig:tectonic_history}). An additional constraint comes from the presence of reliable paleomagnetic poles themselves that are complete with baked contact tests. In a stagnant lid regime, there would not be sufficient heat flow to sustain a geodynamo \citep{Nimmo2000a}. Therefore, the interpretation of stagnant lid for an extended interval implies that Earth should not have had a geomagnetic field that would have enable the preservation of paleomagnetic poles with positive field tests. The record of poles with positive baked contact tests through the Paleoproterozoic and Mesoproterozoic necessitates a geomagnetic field which itself implies operational plate tectonics. This interpretation of a significant persistent geomagnetic field through much of the Proterozoic (with the potential exception of the Ediacaran; \citealp{Bono2019a}) is further bolstered by estimates of paleointensity obtained from mafic dikes from Laurentia (e.g. \citealp{Macouin2006a}) and elsewhere.

The record of these poles also show that there was progressive motion of Laurentia through the Proterozoic (Figs. \ref{fig:Laurentia_paleolatitude} and \ref{fig:Laurentia_reconstructions}). Using data from Laurentia alone, however, it is difficult to ascertain whether this motion is due to plate tectonic motion or rotation of the entire solid Earth through true polar wander. True polar wander can lead to changing position relative to the spin axis even with a stagnant lid. One interval when the Laurentian paleomagnetic record demands that some of the motion is through differential plate tectonics is in the latest Mesoproterozoic. At that time, the pole path is very well-resolved with many high-quality paleomagnetic poles between 1110 and 1070 Ma (Table 2; \ref{fig:Laurentia_poles}). The progression of the poles requires rotation about an Euler pole that is distinct from a great circle path which would result if the motion were solely due to true polar wander \citep{Swanson-Hysell2019a}. These poles constrain rapid motion of Laurentia leading up to collisional orogenesis associated with the Grenvillian orogeny as illustrated in Figure \ref{fig:Grenville_reconstructions}. These data provide strong evidence for differential plate motion at the time and are inconsistent with a stagnant lid. Rather the orogenic cycle of the Mesoproterozoic bears similarity with that of the Paleozoic and reveals Laurentia to have been a central player in the building of amalgamated continents associated with Rodinia and Pangea.

\section{Conclusion}
